---
title: "Lake Empirical Summary"
author: "Luke Winslow"
date: "April 1, 2016"
output: html_document
---

### Summary of empirical lake temperature trends.

---

```{r, echo=FALSE}

if(!exists('lake_slopes')){
  load('../data/temperature_lakes/slope_population.Rdata')
}

metadata = read.csv('../data/temperature_lakes/lake_metadata.csv', as.is=TRUE)

meta_match_i = match(lake_slopes$site_id, metadata$site_id)
lake_slopes$area_m2 = metadata$area_m2[meta_match_i]
lake_slopes$area_log10 = floor(log10(lake_slopes$area_m2))
lake_slopes$zmax    = metadata$zmax[meta_match_i]
lake_slopes$rel_depth = lake_slopes$depth/lake_slopes$zmax
lake_slopes$rel_depth_10 = floor(lake_slopes$rel_depth*10)/10

lake_slopes = lake_slopes[lake_slopes$rel_depth <= 1, ]

  
```

<style type="text/css">
.table {
    width: 60%;
}
</style>


| Metric | Value | 
|--------|-------|
|Number of Lakes| `r length(unique(lake_slopes$site_id))`|
|Overall median trend| `r median(lake_slopes$slopes, na.rm=TRUE)` C/yr|
|Overall near-surface trend (rel_depth <= 0.2)| `r median(subset(lake_slopes, rel_depth <=0.2)$slopes)` C/yr|
|Overall near-bottom trend (rel_depth >= 0.8)| `r median(subset(lake_slopes, rel_depth >=0.8)$slopes)` C/yr|


Depth | Big big lakes (>1 km^2) | Large Lakes (>0.5 km^2) | Small Lakes (<0.5 km^2)
------|-------------|------------| -----------
Shallow (rel_depth < 0.5) | `r median(subset(lake_slopes, rel_depth < 0.5 & area_m2 > 1e6)$slopes)` | `r median(subset(lake_slopes, rel_depth < 0.5 & area_m2 > 0.5e6)$slopes)` | `r median(subset(lake_slopes, rel_depth < 0.5 & area_m2 < 0.5e6)$slopes)`
Deep (rel_depth > 0.5) | `r median(subset(lake_slopes, rel_depth > 0.5 & area_m2 > 1e6)$slopes)` | `r median(subset(lake_slopes, rel_depth > 0.5 & area_m2 > 0.5e6)$slopes)` | `r median(subset(lake_slopes, rel_depth > 0.5 & area_m2 < 0.5e6)$slopes)`
All Lakes | `r median(subset(lake_slopes, rel_depth < 0.5)$slopes)` | `r median(subset(lake_slopes, rel_depth < 0.5)$slopes)` | `r median(subset(lake_slopes, rel_depth < 0.5)$slopes)`




### General lake attributes

---

```{r, echo=FALSE, warning=FALSE}
# library(leaflet)
# 
# color.gradient <- function(x, colors=c("blue", 'white',"red"), colsteps=100) {
#   x[is.na(x)] = median(x, na.rm=TRUE)
#   return( colorRampPalette(colors) (colsteps) [ findInterval(x, seq(min(x),max(x), length.out=colsteps)) ] )
# }
# 
# metadata$secchi[metadata$secchi<=0] = 0.1
# 
# m = leaflet() %>% addTiles() %>% 
#   addCircleMarkers(lng=metadata$lon, lat=metadata$lat, fillOpacity=1,
#                   fillColor=color.gradient(log10(metadata$secchi), colors=c('blue', 'brown')), radius=4, stroke=FALSE, weight=1)#,
#                    #popup=paste("ID:", good_slopes$site_id, '<br>Slope:', good_slopes$V1))
# 
# m

hist(subset(metadata, site_id %in% unique(lake_slopes$site_id))$secchi, breaks=100, xlim=c(0,5), xlab='Avg Secchi by lake (m)', main='Secchi')

hist(log10(subset(metadata, site_id %in% unique(lake_slopes$site_id))$area_m2), breaks=100, xlab='Lake Area as log10(area m^2)', main='Lake Area')


```


### Trends across important gradients

---

```{r, echo=FALSE}


#ddply(lake_slopes, 'rel_depth_10', function(df){
#  median(df$slopes)
#})

boxplot(slopes~rel_depth_10, lake_slopes, ylim=c(-0.2, 0.2), xlab='Relative Depth (depth/zmax)', ylab='Temp Trend (C/yr)')
abline(h=0)

boxplot(slopes~area_log10, lake_slopes, ylim=c(-0.2, 0.2), xlab='Log10(Area) (log10(m^2))', ylab='Temp Trend (C/yr)')
abline(h=0)

```

### Trends split by lake size

---

```{r, echo=FALSE}

boxplot(slopes~rel_depth_10, subset(lake_slopes, area_m2 > 1e6), ylim=c(-0.2, 0.2), 
        xlab='Relative Depth (depth/zmax)', ylab='Temp Trend (C/yr)', main='Large Lakes > 1 km^2')
abline(h=0)

boxplot(slopes~rel_depth_10, subset(lake_slopes, area_m2 < 1e6), ylim=c(-0.2, 0.2), 
        xlab='Relative Depth (depth/zmax)', ylab='Temp Trend (C/yr)', main='Small Lakes < 1 km^2')
abline(h=0)


```

All fairly similar to [Winslow et al. 2015](http://onlinelibrary.wiley.com/doi/10.1002/2014GL062325/full).
Small lakes have muted trends in bottom waters. Surface waters are fairly consistent across large 
and small lakes. 


### Temporal coverage of dataset

```{r, echo=FALSE}

rm(list=ls())

wtemp = read.csv('../data/temperature_lakes/lake_obs.csv', as.is=TRUE)

hist(wtemp$year, xlab='Year', ylab='Number of Observations', main="")

```



